# Influence analysis

This is the interative notebook for [Krishnaraj, P. M., Ankith Mohan, and K. G. Srinivasa. "Performance of procedures for identifying influentials in a social network: prediction of time and memory usage as a function of network properties." Social Network Analysis and Mining 7.1 (2017): 34.](https://link.springer.com/content/pdf/10.1007%2Fs13278-017-0454-1.pdf)

*Set the following parameters:*

```{r}
  input_file = "~/Influence/data/fulwtdata.txt"
  file_sep = " "
  file_header = F
  output_path = "~/Influence/output/"
  set.seed(123)
```

Load the necessary packages

```{r}
  #Install packages if not present
  install <- function(mypkg){
    if (!(mypkg %in% installed.packages())){
      cat("Package not installed, so installing with dependencies... \n Press CTRL C to abort.")
      Sys.sleep(5)
      install.packages(mypkg, dependencies = T)
    }
  }

  #Install packages
  #install(ProNet) No longer supported by CRAN
  install(igraph)
  install(abind)
  install(sna)
  install(ergm.count)
  install(network)
  #install(tcltk) No longer supported by CRAN
  install(tcltk2)
```

### Diagnostics

1. Check if `input_file` exists

```{r}
  if (!file.exists(input_file)){
    stop(input_file, " doesn't exist")
  }
```

2. Check if `output_path` exists, if not create it

```{r}
  if (!dir.exists(output_path)){
    warning(output_path, " doesn't exist")
    cat("Creating", output_path, "...")
    dir.create(output_path)
  }
```

### Helper function

`visualize` function helps in interactive plotting of the graph passed as its parameter

```{r}
  visualise <- function(object = " ", vcolor = "gold", save_file = " "){
    id <- tkplot(object, canvas.width = 1300, canvas.height = 680,
                 vertex.color = vcolor, layout = layout.fruchterman.reingold)
    canvas <- tk_canvas(id)
    tkpostscript(canvas, file = save_file)
    tk_close(id)
  }
```


### Preprocessing

Create a graph from the dataset

```{r}
  input <- read.csv(input_file, sep = file_sep, header = file_header)
  graph <- simplify(graph.data.frame(input, directed = F))
```

### Initialization

```{r}
  percentage <- 0.1
  major_cluster_count <- vector("numeric", 9)
  temporary <- vector("numeric", 9)
  difference <- vector("numeric", 9)
  result <- 0
```

### Sampling (Section 3.2.2 of the paper)

Stopping criterion (Algorithm 4)

```{r}
  stop_index <- 1
  while(percentage < 1){
    length <- vcount(graph) * percentage
    
    # Since ProNet is no longer supported by CRAN, `extraction` function can no
    # longer be used
    # subgraph <- extraction(graph, mode = "sample", sample.number = length)
    
    subgraph <- induced_subgraph(graph, sample(V(graph), length))
    fastgreedy_results <- cluster_fast_greedy(subgraph)
    V(subgraph)$color <- fastgreedy_results$membership
    plot(subgraph, main = paste(percentage * 100, "% sample"),
         vertex.color = V(subgraph)$color, layout = layout.fruchterman.reingold)
    visualise(object = subgraph, vcolor = V(subgraph)$color,
              save_file = paste(output_path, "/sample/sample_", percentage * 100, ".eps", sep = ""))

    #Store only those clusters that have greater than 1% of total number of vertices
    #that were sampled in major_size_fastgreedy_results
    size_fastgreedy_results <- sizes(fastgreedy_results)
    cut_off <- 0.01*length
    major_size_fastgreedy_results <- size_fastgreedy_results[size_fastgreedy_results>cut_off]
    major_cluster_count[stop_index] <- length(major_size_fastgreedy_results)

    #Stopping criterion
    if(stop_index == 1){
      temporary[1] = major_cluster_count[stop_index]
    }else if(stop_index == 2){
      temporary[2] = major_cluster_count[stop_index]
      difference[1] = abs(temporary[2] - temporary[1])
    }else if(stop_index == 3){
      temporary[3] = major_cluster_count[stop_index]
      difference[2] = abs(temporary[3] - temporary[2])
      result = abs(difference[2] - difference[1])
    }else{
      temporary[stop_index] = major_cluster_count[stop_index]
      difference[stop_index-1] = abs(temporary[stop_index] - temporary[stop_index-1])
      if(abs(difference[stop_index-1] - difference[stop_index-2]) <= result){
        result = abs(difference[stop_index-1] - difference[stop_index-2])
        break
      }else{
        result = abs(difference[stop_index-1] - difference[stop_index-2])
      }
    }
    stop_index <- stop_index + 1
    percentage = percentage + 0.1
  }
```

Stabilized graph and clusters

```{r}
  print("Stabilised sampled graph:")
  plot(subgraph, main = paste("Stabilised sample graph"),
         vertex.color = V(subgraph)$color, layout = layout.fruchterman.reingold)
  print("Comunities in the graph:")
  print(fastgreedy_results)
```

Get all major clusters from the stabilized graph

```{r}
  final_cut_off <- 0.01 * vcount(subgraph)
  final <- fastgreedy_results[sizes(fastgreedy_results) > final_cut_off]
  print("Major clusters:")
  print(final)
```

Convert all of these major clusters into matrices

```{r}
  final_matrix <- matrix(nrow = nrow(get.edgelist(graph)), ncol = 2)
  for(final_matrix_cluster in 1:length(names(final))){
    if(final_matrix_cluster != 1){
      final_matrix <- abind(final_matrix, final_matrix, along = 3)
    }
  }

  final_edge <- matrix(NA, ncol=2)
  final_index <- 1
  for(final_names in names(final)){
    noneed <- V(subgraph)[membership(fastgreedy_results) != as.integer(final_names)]
    initial_cluster <- delete_vertices(subgraph, noneed)
    initial_cluster_edge <- get.edgelist(initial_cluster)
    final_matrix[1:nrow(initial_cluster_edge),,final_index] <- initial_cluster_edge
    final_index <- final_index + 1
    if(final_names == names(final)[1]){
      final_edge <- initial_cluster_edge
    }else{
      final_edge <- rbind(final_edge, initial_cluster_edge)
    }
  }
```

Get `not yet`

```{r}
  not_yet_edge <- matrix(NA, ncol = 2)
  not_yet_edge <- rbind(get.edgelist(graph))
  for(final_edge_row in 1:nrow(final_edge)){
    not_yet_edge_row <- 1
    while(not_yet_edge_row <= nrow(not_yet_edge)){
      if((final_edge[final_edge_row,1] == not_yet_edge[not_yet_edge_row,1]) && (final_edge[final_edge_row,2] == not_yet_edge[not_yet_edge_row,2])){
        not_yet_edge <- not_yet_edge[-not_yet_edge_row,]
        break
      }else{
        not_yet_edge_row <- not_yet_edge_row + 1
      }
    }
  }
```

Choose vertices from `not yet` and place them in clusters of highest PageRank or `introverts`

```{r}
  #for every edge in not_yet_edge
  while(nrow(not_yet_edge) > 0){

    #not_yet_vertices has vertices of not_yet_graph
    not_yet_vertices <- V(graph.data.frame(not_yet_edge))

    #all_pageranks is a vector that stores the pagerank of not_yet in each cluster
    all_pageranks <- vector("numeric", length(names(final)))

    #not_yet_random contains randomly selected node from not_yet and removed from not_yet
    not_yet_random <- sample(not_yet_vertices, 1)
    not_yet_random <- names(not_yet_random)
    print("Randomly selected vertex to be clustered:")
    print(not_yet_random)

    #edge_index contains the indices of all the edges corresponding to not_yet_random
    edge_index <- which(not_yet_edge == not_yet_random)
    correct_index <- which(edge_index > nrow(not_yet_edge))
    edge_index[correct_index] <- edge_index[correct_index] - nrow(not_yet_edge)
    edge_index <- unique(edge_index)

    #not_yet_random_edge contains edges corresponding to not_yet_random
    not_yet_random_edge <- matrix(NA, ncol = 2)
    if(NA %in% not_yet_random_edge){
      not_yet_random_edge <- rbind(not_yet_edge[edge_index,])
    }else{
      not_yet_random_edge <- rbind(not_yet_random_edge, not_yet_edge[edge_index,])
    }
    print("Edges corresponding to the randomly selected vertex:")
    print(not_yet_random_edge)

    #remove not_yet_random_edges from not_yet_edge
    not_yet_edge <- not_yet_edge[-edge_index,]

    if(class(not_yet_edge) == "character"){
      not_yet_edge <- t(as.matrix(not_yet_edge))
    }

    #for all clusters in final_matrix
    for(final_matrix_index in 1:length(names(final))){
      for(final_matrix_row in 1:nrow(final_matrix[,,final_matrix_index])){
        if((NA %in% final_matrix[final_matrix_row,1,final_matrix_index]) == TRUE){
          stop_value <- final_matrix_row
          break
        }
      }

      #find the indexed cluster
      nonNA <- stop_value - 1
      cluster_index_edge <- matrix(NA, nrow = nonNA, ncol = 2)
      cluster_index_edge[1:nonNA, 1] <- final_matrix[1:nonNA, 1, final_matrix_index]
      cluster_index_edge[1:nonNA, 2] <- final_matrix[1:nonNA, 2, final_matrix_index]

      #add those edges in not_yet_edge to the indexed cluster if it has a link to a vertex in that cluster
      for(not_yet_index in 1:nrow(not_yet_random_edge)){
        if(((not_yet_random == not_yet_random_edge[not_yet_index,1]) && (not_yet_random_edge[not_yet_index,2] %in% cluster_index_edge) == TRUE) || ((not_yet_random == not_yet_random_edge[not_yet_index,2]) && (not_yet_random_edge[not_yet_index,1] %in% cluster_index_edge) == TRUE)){
          cluster_index_edge <- rbind(cluster_index_edge, not_yet_random_edge[not_yet_index,])
        }
      }

      #find pagerank of not_yet_random in the indexed cluster iff it is present
      #else assign it as zero
      if((not_yet_random %in% cluster_index_edge) == TRUE){
        cluster_index_graph <- graph.data.frame(cluster_index_edge, directed = TRUE)
        cluster_index_pagerank <- page_rank(cluster_index_graph)$vector
        all_pageranks[final_matrix_index] <- cluster_index_pagerank[[not_yet_random]]
      }else{
        all_pageranks[final_matrix_index] <- 0
      }
    }
    print("PageRank of the randomly selected vertex in all the clusters:")
    print(all_pageranks)

    #if not_yet_random cannot be placed in any of the existing clusters then reject all the corresponding edges
    #else add it to the identified cluster
    if(sum(all_pageranks) > 0){
      #find the cluster in which the not_yet element has highest pagerank
      index <- which(max(all_pageranks) == all_pageranks)
      print("Cluster to which the randomly selected vertex has been determined to belong:")
      print(index)

      #find the index of the last edge
      for(final_matrix_row in 1:nrow(final_matrix[,,index])){
        if((NA %in% final_matrix[final_matrix_row,1,index]) == TRUE){
          stop_value <- final_matrix_row
          break
        }
      }

      #find the indexed cluster
      nonNA <- stop_value - 1

      #add_ith_cluster_edge has edges in final_matrix[,,index]
      add_ith_cluster_edge <- matrix(NA, nrow = nonNA, ncol = 2)
      add_ith_cluster_edge[1:nonNA, 1] <- final_matrix[1:nonNA, 1, index]
      add_ith_cluster_edge[1:nonNA, 2] <- final_matrix[1:nonNA, 2, index]

      #add_final_edge contains those edges that are to be added
      add_final_edge <- matrix(NA, ncol = 2)
      for(add_not_yet_index in 1:nrow(not_yet_random_edge)){
        if((not_yet_random_edge[add_not_yet_index,1] %in% add_ith_cluster_edge) == TRUE || (not_yet_random_edge[not_yet_index,2] %in% add_ith_cluster_edge) == TRUE){
          if(NA %in% add_final_edge){
            add_final_edge <- rbind(not_yet_random_edge[add_not_yet_index,])
          }else{
            add_final_edge <- rbind(add_final_edge, not_yet_random_edge[add_not_yet_index,])
          }
        }
      }

      #find start and end index to add edges to final_matrix[,,index] and put add_final_edge there
      start_index <- stop_value
      end_index <- stop_value + nrow(add_final_edge) - 1
      final_matrix[start_index:end_index, 1, index] <- add_final_edge[1:nrow(add_final_edge), 1]
      final_matrix[start_index:end_index, 2, index] <- add_final_edge[1:nrow(add_final_edge), 2]
    }
  }
```

Create necessary folders in `output` directory

```{r}
  create_dir_if_nonexist <- function(output_path, name){
    path = paste(output_path, name, sep = "")
    if (!dir.exists(path)){
      warning("Creating directory ", path, "...\n")
      dir.create(path)
    }
  }

  for (name in c('/abscut/', '/cluster/', '/fixcut/', '/randhist/', '/randplot/', '/sample/', '/screeplot/', '/sdcut/')){
    create_dir_if_nonexist(name)
  } 
```


Convert each of these clusters to graph objects and find influentials in each of them

```{r}
  for(each_cluster in 1:length(names(final))){
    for(final_matrix_row in 1:nrow(final_matrix[,,each_cluster])){
      if((NA %in% final_matrix[final_matrix_row,1,each_cluster]) == TRUE){
        stop_value <- final_matrix_row
        break
      }
    }

    #find the indexed cluster
    nonNA <- stop_value - 1
    real_cluster <- matrix(NA, ncol = 2)
    real_cluster <- rbind(final_matrix[1:nonNA, , each_cluster])
    cluster_graph <- graph.data.frame(real_cluster, directed = TRUE)
    plot(main = paste("Cluster", each_cluster, sep = " "), cluster_graph,
         vertex.color = "gold", layout = layout.fruchterman.reingold)
    visualise(object = cluster_graph,
              save_file = paste(output_path, "/cluster/cluster_", each_cluster, ".eps", sep = ""))

    #get adjacency matrix of the cluster
    cluster_graph_adj <- get.adjacency(cluster_graph, sparse = FALSE)

    #generate screeplot of the dataframe and plot
    in_ties <- sna::degree(cluster_graph_adj, gmode = "digraph", cmode = "indegree")
    names(in_ties) <- names(V(cluster_graph))
    desc_in_ties <- sort(in_ties, decreasing = TRUE)
    plot(desc_in_ties, main = paste("Scree plot of cluster", each_cluster,
                                    sep = " "), xlab = "Actor", ylab = "In-Ties")
    legend('topright', c("Absolute Cut", "Fixed %", "Standard Deviation") ,
           lty=1, col=c('red', 'blue', 'green'), bty='n', cex=1)

    #absolute cut score
    cutoff_abs <- 0.8 * max(desc_in_ties)
    abline(h = cutoff_abs, col = "red")
    abscut <- names(V(cluster_graph)[in_ties > cutoff_abs])

    #Fixed percentage of population
    cutoff_fix <- 0.2 * nrow(cluster_graph_adj)
    abline(v = cutoff_fix, col = "blue")
    fixcut <- names(head(desc_in_ties, cutoff_fix))

    #Standard deviation
    cutoff_sd <- mean(in_ties) + 2 * sd(in_ties)
    abline(h = cutoff_sd, col = "green")
    sdcut <- names(V(cluster_graph)[in_ties > cutoff_sd])

    #save plot as Screeplot<no>.png
    png(file = paste(output_path, "/screeplot/screeplot_", each_cluster, ".png", sep =""),
        height = 1600, width = 1000)
    plot(desc_in_ties, main = paste("Scree plot of cluster", each_cluster,
                                    sep = " "), xlab = "Actor", ylab = "In-Ties")
    legend('topright', c("Absolute Cut", "Fixed %", "Standard Deviation") ,
           lty=1, col=c('red', 'blue', 'green'), bty='n', cex=1)
    abline(h = cutoff_abs, col = "red")
    abline(v = cutoff_fix, col = "blue")
    abline(h = cutoff_sd, col = "green")
    dev.off()

    #color nodes of abscut and plot these iff there are influentials
    if(length(abscut) > 0){
      color <- vector("numeric", length = length(V(cluster_graph)))
      index <- 1
      while(index <= length(abscut)){
        c <- which(names(V(cluster_graph)) == abscut[index])
        color[c] <- 1
        index <- index + 1
      }
      plot(cluster_graph, main = paste("Influentials under Absolute Cut Score Method in cluster", each_cluster, sep = " "),
           layout = layout.fruchterman.reingold, vertex.color = ifelse(color==1, "red", "gold"))
      legend('topright', c("Influentials", "Non-influentials"),
             col=c('red', 'gold'), bty='n', cex=1, pch=19)
      visualise(object = cluster_graph, vcolor = ifelse(color==1, "red", "gold"),
                save_file = paste(output_path, "/abscut/abscut_", each_cluster, ".eps", sep = ""))
    }else{
      print(paste("No influentials have been generated by absolute cut method for cluster", each_cluster, sep = " "))
    }

    #color nodes of fixed% and plot these iff there are influentials
    if(length(fixcut) > 0){
      color <- vector("numeric", length = length(V(cluster_graph)))
      index <- 1
      while(index <= length(fixcut)){
        c <- which(names(V(cluster_graph)) == fixcut[index])
        color[c] <- 1
        index <- index + 1
      }
      plot(cluster_graph, main = paste("Influentials under Fixed Percentage of Population Method in cluster", each_cluster, sep = " "),
           layout = layout.fruchterman.reingold, vertex.color = ifelse(color==1, "dodgerblue", "gold"))
      legend('topright', c("Influentials", "Non-influentials"),
             col=c('dodgerblue', 'gold'), bty='n', cex=1, pch=19)
      visualise(object = cluster_graph, vcolor = ifelse(color==1, "dodgerblue", "gold"),
                save_file = paste(output_path, "/fixcut/fixcut_", each_cluster, ".eps", sep = ""))
    }else{
      print(paste("No influentials have been generated by fixed percentage of population method for cluster", each_cluster, sep = " "))
    }

    #color nodes of sd and plot these iff there are influentials
    if(length(sdcut) > 0){
      color <- vector("numeric", length = length(V(cluster_graph)))
      index <- 1
      while(index <= length(fixcut)){
        c <- which(names(V(cluster_graph)) == fixcut[index])
        color[c] <- 1
        index <- index + 1
      }
      plot(cluster_graph, main = paste("Influentials under Standard Deviation Method in cluster", each_cluster, sep = " "),
           layout = layout.fruchterman.reingold, vertex.color = ifelse(color==1, "green", "gold"))
      legend('topright', c("Influentials", "Non-influentials"),
             col=c('green', 'gold'), bty='n', cex=1, pch=19)
      visualise(object = cluster_graph, vcolor = ifelse(color==1, "green", "gold"),
                save_file = paste(output_path, "/sdcut/sdcut_", each_cluster, ".eps", sep = ""))
    }else{
      print(paste("No influentials have been generated by standard deviation method for cluster", each_cluster, sep = " "))
    }

    #Random Permutation
    #Check the number of vertices in the cluster
    if(length(V(cluster_graph)) >= 5){
      #Fit an ergm model for g1
      g_model <- ergm(cluster_graph_adj ~ edges)

      if(g_model$coef[[1]] != Inf){
        #Conditional simulation on outdegrees for the ergm model
        no_of_sim <- 1000
        g_model.sim <- simulate(g_model, constraints = ~odegrees, nsim = no_of_sim)

        #Store the indegrees of all the simulations in a vector
        in.deg.list <- vector(mode = "numeric", length = 0)
        count <- 0

        for(sim_index in 1:no_of_sim){
          sim_mat <- as.matrix(g_model.sim[[sim_index]])
          sim_in_deg <- sna::degree(sim_mat, gmode = "digraph", cmode = "indegree")
          len <- length(sim_in_deg)

          count <- length(in.deg.list)
          min <- count + 1
          max <- count + len
          in.deg.list[min:max] <- sim_in_deg[1:len]
        }

        #histogram of the indegree alongwith the cut-off
        q <- quantile(in.deg.list, prob = 0.95)
        hist(in.deg.list,
             main = paste("Histogram of Random Permutation of cluster", each_cluster,
                          sep = " "), xlab = "In-degree")
        abline(v = q, col = "navy")
        legend('topright', "alpha = 0.05", lty = 1, col = "navy",
               bty = 'n', cex = 1)

        #save plot as RandomHist<no>.png
        png(file = paste(output_path, "/randhist/randhist", each_cluster, ".png", sep = ""),
            width = 1600, height = 1000)
        hist(in.deg.list,
             main = paste("Histogram of Random Permutation of cluster", each_cluster,
                          sep = " "), xlab = "In-degree")
        abline(v = q, col = "navy")
        legend('topright', "alpha = 0.05", lty = 1, col = "navy",
               bty = 'n', cex = 1)
        dev.off()

        cutoff <- q[[1]]
        data.in.deg <- sna::degree(cluster_graph_adj, gmode = "digraph", cmode = "indegree")

        names <- names(V(cluster_graph))
        names.list <- vector(mode = "character", length = 0)
        index.list <- vector(mode = "numeric", length = 0)
        x <- 1
        for(i in 1:length(data.in.deg)){
          if(data.in.deg[i] >= cutoff){
            names.list[x] <- names(V(cluster_graph)[i])
            index.list[x] <- i
            x = x + 1
          }
        }

        #color nodes of rand and plot these
        color <- vector("numeric", length = length(V(cluster_graph)))
        index <- 1
        while(index <= length(names.list)){
          c <- which(names(V(cluster_graph)) == names.list[index])
          color[c] <- 1
          index <- index + 1
        }
        plot(cluster_graph, main = paste("Influentials under Random Permutation Method in cluster", each_cluster, sep = " "),
             layout = layout.fruchterman.reingold,
             vertex.color = ifelse(color==1, "chocolate", "gold"))
        legend('topright', c("Influentials", "Non-influentials"),
               col=c('chocolate', 'gold'), bty='n', cex=1, pch=19)
        visualise(object = cluster_graph, vcolor = ifelse(color==1, "chocolate", "gold"),
                  save_file = paste(output_path, "/randplot/randplot_", each_cluster, ".eps", sep = ""))
      }else{
        print(paste("The MLE coefficient is Inf. Cluster", each_cluster,
                    "is saturated resulting in no influentials.", sep = " "))
      }
    }else{
      print(paste("The cluster", each_cluster, "has", length(names(V(cluster_graph))), "nodes.",
                  sep = " "))
    }
  }
```

